# **Assignment 3**

### **Step 1: Install and Load Packages**

```{r}
# Install the necessary packages if not already installed
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("corrplot")
install.packages("dplyr")

# Load the libraries
library(conflicted)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(dplyr)
```

### **Step 2: Load the Dataset**

```{r}
# Load the dataset
supermarket_sales <- read.csv("C:\\Users\\User\\OneDrive - International Islamic University Malaysia\\Documents\\Mater courses\\Sem 1\\TTTR6124 STATISTICAL METHODS FOR DATA ANALYTICS\\Assignments\\supermarket_sales.csv")

# View the first few rows of the dataset
head(supermarket_sales)
```

### Step 3: Data Analysis

#### Descriptive Analysis

1.  Summary Statistics

2.  Sales Distribution by Branch, City, Product Line, and Customer Demographics

```{r}
# Summary statistics for sales and quantities
summary(supermarket_sales$Total)
summary(supermarket_sales$Quantity)

# Sales distribution by branch
ggplot(supermarket_sales, aes(x = Branch, y = Total)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Sales Distribution by Branch")

# Sales distribution by city
ggplot(supermarket_sales, aes(x = City, y = Total)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Sales Distribution by City")

# Sales distribution by product line
ggplot(supermarket_sales, aes(x = Product.line, y = Total)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Sales Distribution by Product Line")

# Sales distribution by customer demographics
ggplot(supermarket_sales, aes(x = Gender, y = Total)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Sales Distribution by Gender")

# Sales distribution by customer type
ggplot(supermarket_sales, aes(x = Customer.type, y = Total)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Sales Distribution by Customer Type")
```

#### Inferential Analysis

1.  ANOVA to Compare Sales Between Different Branches

2.  Regression Analysis

```{r}
# Check the column names in the dataset
colnames(supermarket_sales)
```

```{r}
# ANOVA to compare sales between branches
anova_result <- aov(Total ~ Branch, data = supermarket_sales)
summary(anova_result)

# ANOVA to compare sales among product lines
anova_result_product <- aov(Total ~ Product.line, data = supermarket_sales)
summary(anova_result_product)
```

```{r}
# Simple Linear Regression to explore the relationship between sales and other variables
regression_result <- lm(Total ~ Quantity + Unit.price + Tax.5. + cogs + gross.income, data = supermarket_sales)

# Summary of the regression model
summary(regression_result)

# Residuals Plot
ggplot(data = supermarket_sales, aes(x = predict(regression_result), y = residuals(regression_result))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  ggtitle("Residuals vs Fitted Plot") +
  xlab("Fitted Values") +
  ylab("Residuals")

# Predicted vs Actual Plot
ggplot(data = supermarket_sales, aes(x = predict(regression_result), y = Total)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_minimal() +
  ggtitle("Predicted vs Actual Plot") +
  xlab("Predicted Values") +
  ylab("Actual Values")
```

#### Visualization

1.  Histogram of Total Sales

2.  Bar Chart of Sales by Branch

3.  Line Chart of Sales Over Time

4.  Correlation Heatmap

```{r}
# Histogram of Total Sales
ggplot(supermarket_sales, aes(x = Total)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black", alpha = 0.7) +
  theme_minimal() +
  ggtitle("Histogram of Total Sales") +
  xlab("Total Sales") +
  ylab("Frequency")

# Bar chart of Sales by Branch
ggplot(supermarket_sales, aes(x = Branch, y = Total, fill = Branch)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  ggtitle("Total Sales by Branch") +
  xlab("Branch") +
  ylab("Total Sales")

# Convert Date column to Date type if not already in that format
supermarket_sales$Date <- as.Date(supermarket_sales$Date, format = "%m/%d/%Y")

# Line chart of Sales Over Time
ggplot(supermarket_sales, aes(x = Date, y = Total)) +
  geom_line(color = "blue") +
  theme_minimal() +
  ggtitle("Sales Over Time") +
  xlab("Date") +
  ylab("Total Sales")

# Select numeric columns for correlation matrix
numeric_columns <- supermarket_sales %>% select(where(is.numeric))

# Remove columns with zero standard deviation
numeric_columns <- numeric_columns %>% select_if(~ sd(.) != 0)

# Calculate the correlation matrix
correlation_matrix <- cor(numeric_columns, use = "complete.obs")

# Plot the correlation heatmap
corrplot(correlation_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         addCoef.col = "black", number.cex = 0.7, 
         col = colorRampPalette(c("red", "white", "blue"))(200))
```
